---
title: "Carpentries around the world"
output: html_document
---

n.b. not an actual dataset as part of R4DS Tidy Tuesday

Key question: how does New Zealand compare to other countries in terms of workshops


```{r setup, inlcude = FALSE}
library(tidyverse)
# library(httr)
library(rvest)
```


Grab the info from The Carpentries about past workshops
```{r}
url <- "http://static.carpentries.org/past_workshops/"
imgs <- read_html(url) %>% html_node("table") %>% html_nodes("img")
flags <-  tibble(class = xml_attr(imgs, "class"), title = xml_attr(imgs, "title"), alt = xml_attr(imgs, "alt")) %>% filter(str_detect(alt, "logo", negate = TRUE))


country_n_workshops <- flags %>% group_by(title) %>% tally()
```



Grab the table from wikipedia that can provide the population information
```{r}
wiki_pop_url <- "https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population"
wiki_pop_page <- read_html(wiki_pop_url) %>% html_node("table")
wiki_pop_table <- wiki_pop_page %>% html_table() %>% set_names(nm = janitor::make_clean_names(names(.)))

# clean up some country names and numeric columns
wiki_pop_table_cleaned <- wiki_pop_table %>% 
  mutate(
    country_or_dependent_territory = str_remove(country_or_dependent_territory, "\\[Note [0-9]+\\]"),
    population = as.numeric(str_remove_all(population, ",")),
    percent_of_worldpopulation = as.numeric(str_remove_all(percent_of_worldpopulation, "%"))
  )
```

Grab the table from wikipedia to link two letter country codes to the population name
```{r}
wiki_country_code_url <- "https://en.wikipedia.org/wiki/ISO_3166-1"

# read the page and extract the table elements (the second table is what we want)
wiki_country_code_table <-read_html(wiki_country_code_url) %>% html_nodes("table") %>% magrittr::extract2(2) %>% html_table(fill = TRUE) %>% set_names(nm = janitor::make_clean_names(names(.)))

# clean some country names
wiki_country_code_table_cleaned <- wiki_country_code_table %>%
  mutate(english_short_name_using_title_case = case_when(
    english_short_name_using_title_case == "United States of America" ~ "United States" , 
    english_short_name_using_title_case == "United Kingdom of Great Britain and Northern Ireland" ~ "United Kingdom",
    TRUE ~ english_short_name_using_title_case)
  )
```


Join the data together
```{r}
country_data <- wiki_country_code_table_cleaned %>% 
  full_join(wiki_pop_table_cleaned, by = c("english_short_name_using_title_case"= "country_or_dependent_territory")) %>% 
  full_join(country_n_workshops, by = c("alpha_2_code"= "title"))
```

Calculate per capita workshops and display table arranged descending by per capita

n.b. need to sort out matching of country names since US and UK (notibly) aren't getting joined correctly
```{r}
per_capita <- country_data %>% 
  filter(!is.na(n)) %>% 
  mutate(perCapita = n / population) %>% 
  select(english_short_name_using_title_case, alpha_2_code, population, n, perCapita) %>% 
  arrange(desc(perCapita))

per_capita %>% knitr::kable()
```


New Zealand hasn't run the most workshops but does seem to have run the most on a per capita basis

```{r}
per_capita %>% arrange(desc(n)) %>% slice(1:20)%>% mutate(alpha_2_code = forcats::fct_reorder(alpha_2_code, .x = n, mean), highlight = ifelse(alpha_2_code  == "NZ", TRUE, FALSE)) %>% ggplot(., aes(x = alpha_2_code, y = n, fill = highlight)) + geom_col() + theme_bw()
```

Arrange per capita by number of workshops run for top 20 countries (based on number of workshops)
```{r}
per_capita %>% 
  arrange(desc(n)) %>% 
  slice(1:20) %>% 
  mutate(alpha_2_code = forcats::fct_reorder(alpha_2_code, .x = n, mean), 
         highlight = ifelse(alpha_2_code  == "NZ", TRUE, FALSE)) %>% 
  ggplot(., aes(x = alpha_2_code, y = perCapita, fill = highlight)) + geom_col() + theme_bw()
```


Top 20 countries by per capita
```{r}
per_capita %>% arrange(desc(perCapita)) %>% slice(1:20)%>% mutate(alpha_2_code = forcats::fct_reorder(alpha_2_code, .x = perCapita, mean), nz_best = ifelse(alpha_2_code  == "NZ", TRUE, FALSE)) %>% ggplot(., aes(x = alpha_2_code, y = perCapita, fill = nz_best)) + geom_col() + theme_bw() + labs(title = "NZ has best per capitia rate of workshops")
```

A fun but not particularly useful statistic since it takes much less effort to influence per capita in small countries than in large population countries.


```{r}
map.world <- map_data('world') %>%
  filter(region != "Antarctica")

per_cap_world <- per_capita %>% mutate(entity = case_when(
    str_detect(english_short_name_using_title_case, "United States") ~ "USA",
    str_detect(english_short_name_using_title_case, "United Kingdom") ~ "UK",
    TRUE ~ english_short_name_using_title_case
))

country_join <- left_join(map.world, per_cap_world, by = c('region' = 'entity')) %>% mutate(perCapita = ifelse(is.na(perCapita), 0, perCapita))



ggplot(data = country_join, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = perCapita), colour = "darkgrey") +
  scale_fill_continuous(high = "yellow", low = "blue", breaks = c(seq( 1e-5, 1e-7, length.out = 7  ),0)) +
  labs(title = "",
       subtitle = "",
       x = "",
       y = "",
       caption = "") +
  guides(
    fill = guide_legend(title = "Workshops Per Capita")) + 
  coord_fixed(ratio = 1.3) + theme_bw()

```

We could even look what proportion of all workshops were taught in particular countries
```{r}
per_capita %>% mutate(prop_workshops = n/sum(n)) %>% arrange(desc(prop_workshops)) %>% slice(1:20)%>% mutate(alpha_2_code = forcats::fct_reorder(alpha_2_code, .x = prop_workshops, max), NZ = ifelse(alpha_2_code  == "NZ", TRUE, FALSE)) %>% ggplot(., aes(x = alpha_2_code, y = prop_workshops, fill = NZ)) + geom_col() + theme_bw() + labs(title = "Proportion of workshops taught by country")
  
```

NZ still doing well but not as favourably.