---
title: "Carpentries around the world"
output: html_document
---

n.b. not an actual dataset as part of R4DS Tidy Tuesday

Key question: how does New Zealand compare to other countries in terms of workshops


```{r setup, inlcude = FALSE}
library(tidyverse)
# library(httr)
library(rvest)
```


Grab the info from The Carpentries about past workshops
```{r}
url <- "http://static.carpentries.org/past_workshops/"
imgs <- read_html(url) %>% html_node("table") %>% html_nodes("img")
flags <-  tibble(class = xml_attr(imgs, "class"), title = xml_attr(imgs, "title"), alt = xml_attr(imgs, "alt")) %>% filter(str_detect(alt, "logo", negate = TRUE))


country_n_workshops <- flags %>% group_by(title) %>% tally()
```



Grab the table from wikipedia that can provide the population information
```{r}
wiki_pop_url <- "https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population"
wiki_pop_page <- read_html(wiki_pop_url) %>% html_node("table")
wiki_pop_table <- wiki_pop_page %>% html_table() %>% set_names(nm = janitor::make_clean_names(names(.)))
```

Grab the table from wikipedia to link two letter country codes to the population name
```{r}
wiki_country_code_url <- "https://en.wikipedia.org/wiki/ISO_3166-1"

# read the page and extract the table elements (the second table is what we want)
wiki_country_code_table <-read_html(wiki_country_code_url) %>% html_nodes("table") %>% magrittr::extract2(2) %>% html_table(fill = TRUE) %>% set_names(nm = janitor::make_clean_names(names(.)))
```


Join the data together
```{r}
country_data <- wiki_country_code_table %>% full_join(wiki_pop_table, by = c("english_short_name_using_title_case"= "country_or_dependent_territory")) %>% full_join(country_n_workshops, by = c("alpha_2_code"= "title"))
```

Calculate per capita workshops and display table arranged descending by per capita

n.b. need to sort out matching of country names since US and UK (notibly) aren't getting joined correctly
```{r}
per_capita <- country_data %>% filter(!is.na(n)) %>% mutate(population = as.numeric(str_remove_all(population, ",")))%>% mutate(perCapita = n / population) %>% select(english_short_name_using_title_case, alpha_2_code, population, n, perCapita) %>% arrange(desc(perCapita))

per_capita %>% knitr::kable()
```


New Zealand hasn't run the most workshops but does seem to have run the most on a per capita basis
